\textbf{\sc Deliberative Autonomy for Marine Robotics}

### 2012-02-02 trunk/arch.tex
A lot of text added to the closing section of TREX : need to be massaged but I think the ideas are already in there

\subsubsection{Intertwining synchronization and execution }
\label{sec:arch:intertwine}

>>>>

\subsubsection{Intertwining synchronization and execution}
\label{sec:arch:intertwine}

A core aspect of this reactor is that it takes advantage of the
specific sequencing enforced by the T-REX architecture along with the
fact that only one $plan$ structure is maintained in the reactor and
shared between synchronization and planning to allow these two processes
to not only be interleaved but impact each other through the way they
manipulate this plan.

The most obvious case is how synchronization allow propagate of {\em
  external} observations as a new tick occur in the plan. In or
presentation of synchronization, we isolated the problem by stating
that we will assume at this point that the reactorhas no future goal
or more accurateely no plan to enforce at this stage. By doing so we
were able to develop the way we augment yhe europpa solver to be able
to propagete {\em external} observations in order to identify the
reactor {\em internal} state for the same tick. By doing so we were
able to focus on synchronization as a state evluation process that can
be resolved within the europa framework with few extensions. simliarly
we presented deliberation at this fairly high level not necessarily
discussing how the disruptive nature of synchronization in this
process could impact the way this planning process will evolve.

We first anlyse the case during which synchronization occurs after the
planning has vcompleted (ie the last seep of deliberation resulted into
a plan that is considered as complete for the current look-ahead). As
synchronization starts with  the plan produced this will gave a frame
for the search process. Therefore this deliberation will not only be a
model-based estimation uniquely based on the {\em external}
observations but will also include the reactor intentions as described
by the plan produced. Consider for example the Shopping Agent model
described during the europa presentation and that the
location of the \texttt{agent} is an {\em external} timeline. For
example as, the reactor add the goal to have \texttt{Own} milk and the
current {\em external} states indicated that it was \texttt{At(home)} it
produced the plan in Fig. \ref{fig:shop:exec0}. The  {\em external}
  part of the plan -- namely the tokens \texttt{Going(home,
    superMarket)} and \texttt{At(SuperMarket)} -- and integrated as
  goals for the reactor managing this specific state varaiable {\em
    internally}.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0}
  \caption{Shopping example : plan produced before synchronization. 
    Tokens in red indicate observations, 
    tokens in blue the goal that this plan attempt to solve. Arrows
    indicate temporal constraints.}
  \label{fig:shop:exec0}
\end{figure}

Now, consider that at the next tick $\tau+1$ the reactor managing the
agent location was able to change the location state varaibale to 
the \texttt{Going}. T-REX informs our reactor of this new observation
which is added in the plan structure as a fact -- starting exactly at
$\tau+1$ and lastign for an unlknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization the solver can identify the similarity between this
new observation oand the next token in the existing plan. Which mean
that instead of making a blind exhaustive search of all the possible
implication the synchronization process starts its execution with the
assumption that these new observations are consitent with the current
plan maintained. As a result at the end of synchronization we will
have the plan depicted in Fig. \ref{fig:shop:exec1}. Which just
propagated the information provided by the new observation  in the
exisiting plan (restricting in turn the start time of 
\texttt{Going} and by propagtion of the constraints \texttt{Go} to  be
{\em exactly} $\tau+1$). While focusing uniquely on the execution
frontier ($\tau+1$ in this figure), the europa solver took benefit of
the exisiting plan to evaluate a solution that asumes that the
observation is consistent with what was decided  inside this
reactor.

This choice is directed not only by some natural heurisitc
choices made in the solver but even more enforced in general by the
heuristic order of the decision point reolution choices for our {\em
  internla} state flaw presented in section \ref{sec:arch:sync}
{\em\color{red} TODO: need to rework the presentation of thee choices
  so I can refer to them more directly -- algorithm or pseudo theorem
  should do the trick}. Should such a flaw be revealed during
synchronization -- meaning that the current state of an {\em internal}
state varaibale is not fully grounded -- the sequencing of the choices
will atempt first a conservative choice in term of maintaining
previous state, then an optimisitic choice in term of the
execution/advance in our current plan to finally attempt other choices
that were not predicted by the model. The two first possible choices
are strongly directed by the exisiting plan which help direct the
search within this frame. This often allow to have a synchronization
that is much more directed and i, in nominal situation such as the one
we showed resolve the deliberation in few steps (often without any
backtrack in the search).

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t1}
  \caption{Shopping example : Result of synchronization after the {\em
      Going} observation was received by the reactor with thew plan
    showed in Fig. \ref{fig:shop:exec0}.}
  \label{fig:shop:exec1}
\end{figure}

The synchronization process is also in this process supporting fully
the tracking of the plan execution as it specifies tokens parameters
such as the time points and allow this to propagate through the plan
in order to idenitfy if the plan remains valid -- as in our example --
or that it cannot be executed as it is. The later  is often identified
by the synchronization failing to find a consistent solution. for
example dshould the \texttt{Going} observation we receive be to the 
\texttt{hardwareStore}, this would break the current plan of our
reactor as we cannot go simultaneously at 2 different places as
illustrated in Fig. \ref{fig:shop:relax}-1. as synchronization is not
possible in this context our strategy is to fully relax all decisions
made by the previous plan steps by just keeping n the plan all the
past observations in both {\em external} and {\em internal} state
variables and the exsiting {\em internal} goals received by this
reactor (\texttt{Own(Milk)} in our example) resulting on the plrtial
plan sghowned in Fig. \ref{fig:shop:relax}-2. In this process the
reactor inform the agent that all {\em external} goals it requested
for the future are note valid anymore which often result in the ownere
of htese timelines to not have to maintain this goal anynmore.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax}  
  \caption{Illustration of a conflict suring synchronization and the resulting relaxed plan for recovery}
  \label{fig:shop:relax}
\end{figure}

As the plan is not anymore in the way the synchronization can resume 
and find a solution as we initial lly described. The newly pending goals will enforce  at
the next \texttt{step} to resumed deliberation on this new fully
relaxed partial plan. By the simple use of synchronization the reactor 
was not only able to identify that its plan was not executable anymore 
but recover from it while offering a new incomplete partial plan to be
resolved during the next deliberation \texttt{steps}. This on its own
ogfer the core functionality one should except to close the loop
between planning and execution in the sense that synchronization plays
the role of what is typically named as the {\em Executive} in other
architectures {\em\color{red} references to diverse archis, 3-tiered,
  laas, claraty, B Wiliams, and many more}.

In our case though the interaction between planning and execution
does not stop at this level. A core point of the overal schedulling of
both {\em synchronization} and {\em planning} steps resides on our
architecture enforsincing that the synchroniuzation process can occur
in between any step of the plan. This lead us to the fact that not
only synchronization is imnfluenced by the outcome of planning but
this influence goes also in the alternate direction. Specifically when
synchronuiation occuers before we found a complete plan, it will
inject as we described new facts (the {\em external} observations and
resulting {\em internal} state) in the plan database. This
informatiuon will be present when the next planning \texttt{step} will
resume. Therefore they are not intergal part of the plan resolution
(generating potentially new flaws for the planner or contributing to
resolve previously exisiting flaws). The planning process is therefore
perturbed internallyand fully informed on how the world evolves as it
is planning.  This aspect trueluy contrast to other approaches we have
idenitifed in the literature which avoid to perturb the planner during
it search as it is not compatible with the ``off-line planning'' assumption.

By our design choices -- both at the architecture level and the integration of
the europa framework for embedded deliberation -- we were able to
implement an architecture that avoid this limitation while remaining in
a formal frame that still allow for deliberation and model-based
agent control. We believe that this tighter integration of planning
and execution allow the system to ba more accute of its environament
alowing better informaed decision and tighter reaction. This is
possible while still integrating information such as a long latency
for the reactors as long as one can ensure that synchronization of al
the reactors and deliberation steps can be done in a time that is
resonably small in comparison of the agent tick duration.
